---
title: "MATH1005 Lab8 Worksheet"
subtitle: "Chance variability (Box Model)"
author: "© University of Sydney 2019"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab8 {-}

**Learning Outcome** 
LO6 Use the box model to describe chance and chance variability, including sample surveys and the central limit theorem. 

At the end of this Lab, you should be able to:

- Draw a box model to model the **sum** of random draws (sample) from a box (population).
- Calculate the mean and SD of the box.
- Calculate the EV and SE of the **sum** of the draws from the box.
- Calculate Normal probabilities for the **sum** of draws from the box. 

Note: You also need to know how to model the **mean** of draws from the population.

**R commands** 
`multicon` (package),`mean`, `popsd`

# <i class="fa fa-th-list"></i> Summary {-}

The **box model** models the sum (or mean) of $n$ random draws from a population. The population is the "box", and the collection of draws forms the "sample".

(1) The Box

As the box represents a population, we use `popsd(box)`, where `popsd` requires the `multicon` package.

|Focus|By hand|In R|Symbol (ext)|
|--------------|-----|--|-------|
|Mean of the Box|Calculate the the average| `mean(box)`|$\mu$ |
|SD of the Box|Calculate the population sd|  `popsd(box)`|$\sigma$ |

(2) The Sample

We can focus on either the Sum (or Mean) of the sample.

|Focus|Expected Value (EV)|Standard Error (SE)|
|----|----|----|
|Sum of the Sample| $n$ $\times$ Mean of the Box |$\sqrt{n}$ $\times$ SD of the Box |
|Mean of the Sample| Mean of the Box | SD of the Box / $\sqrt{n}$|

# StatQuiz

# StatBook Question 
p277 Set A: Q3,5, p280-281 Set B: Q2,6, p284 Set C: Q2,3, p293-4 Set B: Q1,4, p296-7 Set C: Q2,5, p303-4 Set E: Q2,6,9

# Simple Box Model 

This is more a "pen and paper" lab.

For each calculation, work out the answer by hand, and then check your answers in R.

Model the following context: A box contains 1,2,3,4,5 and we draw 25 times with replacement. Focus on the **sum** of the draws.

## Draw the box model.

Note: On the Carslaw computers, the `DiagrammeR` package may not work.

```{r, fig.height=3, eval=F}
library("DiagrammeR")
par(mfrow=c(1,4))
grViz(" 
  digraph CFA {

  # All
  node [fontname = Helvetica, fontcolor = White]

    # Box
    node [shape = box, style=filled, color=SteelBlue4, width=2 label='1 2 3 4 5'][fillcolor = SteelBlue4]
    a ; 

    # Sample
    node [shape = circle, style=filled, color=SteelBlue3, width=0.5, label='   '][fillcolor = SteelBlue3]
    b ; 

    # Draws
    a -> b [label = '25 draws  ',fontsize=8]
    b -> a  [color=grey,arrowsize = 0.5]
  }

")
detach(package:DiagrammeR)
```

## Calculate the Mean and SD of the box

```{r, eval=F}
box=c(1,2,3,4,5)
mean(box)
library(multicon)
popsd(box) 
```

## Calculate the EV and SE of the sum of draws from the box

```{r, eval=F}
25* mean(box)
sqrt(25) * popsd(box)
```

## Find the chance that the sum of the draws is between  70 and 80 (inclusive). 

- See the picture on p312 (Q1) of your textbook. From looking at the histogram, what would you predict the area to be? 
    
- Extension (for the uber coders). Draw a line underneath representing the standard units.

```{r, eval=F}
ev = 25* mean(box)
library(multicon)
se = sqrt(25) * popsd(box)
f <- function(x) dnorm(x, ev, se)
plot(f, ev-3*se, ev+3*se)
axis(1, line=0, at = -3:3, labels = c("-3s", "-2s", "-1s", "mean", "1s", "2s", "3s"))

curve(f, ev-3*se, ev+3*se, lwd=2, axes = FALSE, xlab = "", ylab = "")
axis(1, line=0, at = c(ev-3*se, ev-2*se, ev-1*se,ev,ev+1*se,ev+2*se,ev+3*se), labels = c(round(ev-3*se,2),round(ev-2*se,2),round(ev-1*se,2),round(ev,2),round(ev+1*se,2),round(ev+2*se,2),round(ev+3*se,2)))
axis(1, line=2, at = c(ev-3*se, ev-2*se, ev-1*se,ev,ev+1*se,ev+2*se,ev+3*se), labels = c("-3s", "-2s", "-1s", "mean", "1s", "2s", "3s"))
```

- Estimate the area using a Normal curve.

```{r, eval=F}
pnorm(80,ev,se)-pnorm(70,ev,se)
```
```{r pressure, echo=FALSE}
plot(pressure)



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
