---
title: "MATH1005 Lab4 Worksheet"
subtitle: "Numerical Summaries"
author: "© University of Sydney 2019"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab4 {-}

**Learning Outcome**  
LO3 Produce, interpret and compare graphical and numerical summaries in R.

At the end of this Lab, you should be able to:

- Choose what numerical summaries are appropriate for given variables, and interpret the output.
- Produce numerical summaries with R (mean, median, SD, IQR).
- Write a concise and precise executive summary of your findings.

**R commands** 
`class`, `hist`, `barplot`,`boxplot`

# StatQuiz

# StatBook Question 
p34 Set A: Q4,5,6; p46 Set E: Q1,2

# <i class="fas fa-home"></i> Sydney Property Market 

The Newtown property prices dataset comes from data scraped by [domain](https://www.domain.com.au/). Who might be interested in this data?

```{r}
property = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/2018S2/data/NewtownJune2017.csv")
```

## Executive Summary

I discovered ...   [Fill this out at the end of your explorations.]

## Initial Data Analysis (IDA)

- Produce a snapshot of the data. What variables do you think would most affect the sale price of property?
```{r}
str(property)
```

## Research Questions

### What were the cheapest properties? 

- Here we consider **1 quantitative variable**: *price* of the properties.

- First, we could just sort by the sold price.

```{r}
# Sort all prices
sold = property$Sold
sort(sold)
``` 

- It would nice to list the details about the cheapest properties. Try the following code.
```{r}
# Cheapest
attach(property)   # The database is attached to the R search path, so variables in the database can be accessed by simply giving their names.
head(property[order(Sold), ][c(1, 6, 7)], 10)
```  

- Go through this last command carefully, noticing what it does.

- Did you expect these properties to be the cheapest?

- Summarise your findings.

### What were the most expensive properties? 

- Now focus on the most expensive properties.

```{r}
# Most expensive
head(property[order(Sold, decreasing =T), ][c(1, 6, 7)], 10)
```   

- Did you expect these properties to be the most expensive?

- Summarise your findings.
    
### What was the average price for a 2 bedroom property? 
  
- Here we consider **1 quantitative variable**: *price* of the 2B properties.

- First we need to filter down to just the 2 bedroom properties.
   
```{r}
# Select 2 bedroom properties
property1 = property$Sold[property$Bedrooms=="2"]
```

- Now consider the mean and median. Are they different to each other here? Why or why not?

```{r}
# Find the mean and median
mean(property1)
median(property1)
```  
    
- What were the standard deviation (SD) and interquartile range (IQR) for the 2 bedroom properties? What information does this give us?

```{r}
# Find the SD and IQR
sd(property1)
IQR(property1)
```  

- Summarise your findings.

###  Does the number of carspots seem to affect the price? 

```{r}
mean(property$Sold[property$Carspots=="0"])
mean(property$Sold[property$Carspots=="1"])
mean(property$Sold[property$Carspots=="2"])
```  

- Postulate reasons for your conclusions.

### Assuming commission fees are around 2%, which agency earnt the most over that period? (Ext)

```{r}
# Example of extracting Agent sales
sum(property$Sold[property$Agent=="RayWhite"])*0.02
```  
    
### Explore your own research question    
    
    